---
title: "x_tab() summary"
author: "Hua"
format: html
editor: visual
---

# !!! This is an extension of the previous data pre-processing steps and highlights. Please ensure that you have run all the code chunks from the VectorAtlas-Hua.qmd file, or that you have saved the variables in your environment.

\
\
\

## Variables used to identify distinct occurrences:

```{r}
colnames(ref_data1) # All variables were used for comparison in the format: !(%in%), except "sources".
```

## Different combined data indicating the number of distinct occurrences gradually added from each source

```{r}
nrow(ref_data1) # Cleaned data from Malaria Atlas in R
nrow(ref_data1_3a) # Cleaned combined data from Malaria Atlas in R and RGBIF
nrow(ref_data1_3a_3b) # Cleaned combined data from Malaria Atlas in R, RGBIF, and GBIF Website
nrow(ref_data1_3a_3b_4) # Cleaned combined data from all sources
```

## To view the number of occurrences slowly added from the newly cleaned data source:

```{r}
xtabs(~species,data=ref_data1) # refer to original Malaria Atlas in R (Ref dataset)
xtabs(~species,data=ref_data1_3a) # refer to Combined-1,3a. in Excel table
xtabs(~species,data=ref_data1_3a_3b) # refer to Combined-1,3a,3b. in Excel table
xtabs(~species,data=ref_data1_3a_3b_4) # refer to Combined-1,3a,3b,4 "Total". in Excel table
```

## This indicates the number of occurrences for each recorded species, referring to the total number of occurrences in the complete combined data set

```{r}
xtabs(~species, data=ref_data1_3a_3b_4) # from complete data
```

## This indicates the number of distinct occurrences extracted from each source, as referenced in the vRef column of the Excel table

```{r}
xtabs(~species+sources, data=ref_data1_3a_3b_4) # from complete data
```
